Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests

### 2018-11-14 paper/sections/methodology.tex
Paul's comments included
wednesday modifications

The primary aim of this research is to enable us to collect a series of probe requests and process them into an usable pedestrian footfall count. 
We do this by using a Wi-Fi receiver to collect probe requests broadcasted by mobile devices, filtering out the background noise and aggregating them based on the device that generated them.
We begin by looking at the characteristics of probe requests in detail, device a methodology to collect these probe requests in public areas, examine the systemic biases and uncertainties in the data collection method and device data processing methods to overcome these challanges.
Finally we compare the processed footfall counts to the ground truth recorded by primary surveys.

Probe requests are a special type of management packets broadcast by Wi-Fi enabled devices as part of the various functions such as scanning for available access points (AP), quick geo-location by triangulation based known APs, etc.
These are broadcast by all Wi-Fi enabled devices regardless of the manufacturer, type or model of the devices though there is some variation on the frequency and the information transmitted through them.
In some cases, such as Android devices, these are broadcast even when the Wi-Fi functionality has been turned off by the user.
Thus these signals can be used to reliably identify the presence of Wi-Fi enabled mobile devices.

Being a first step of connection initiated by the mobile device, these packets have information regarding the characteristics of the mobile device itself. Some of the key information we can infer from these requests are,
\begin{enumerate}
	\item \textbf{Media Access Control (MAC) address} which is an unique identifier for the wireless hardware of the mobile device,
	\item \textbf{Sequence number} of the request for the mobile device to keep track of the responses,
	\item \textbf{Timestamp} at which the request was received by the AP,
	\item Total \textbf{length} of the request in number of bits, and 
	\item The \textbf{strength of the signal} which transmitted the request.
\end{enumerate}
The MAC address is the primary unique identifier for the mobile device.
It has two parts, first part is called an Organisation Unique Identifier (OUI) which gives information about the manufacturer of the device and the second is unique to the device. The MAC address can be randomised (hence non unique) and is marked as such. Though sequence number and length of the packet are not strictly unique, we hypothesize that we can use them to estimate unique devices.

Data collection was done with the help of custom sensors built from modifing the Smart street sensor \citep{sss2016} hardware and updating them with custom software.
The sensor is essentially a Raspberry Pi connected with Wi-Fi and 3G antennae.
It keeps the Wi-Fi module in `Monitor' mode and uses the open source software - wireshark cite to passively collect all packets sent to `broadcast', marked with type - `management' and subtype - `probe requests'.
The MAC address in these probe requests is anonymised using a cryptographic hashing algorithm and transmitted through 3G connection to a central database via web-sockets protocol, where it is stored in a PostgreSQL database for further analysis.
A overall schematic of the data collection process is shown in Figure \ref{datacollection_schematic}.
The ground truth on number of pedestrian footfall was recorded using a purpose built Android application cite.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{images/datacollection_schematic.jpeg}
	\caption{Schematic diagram showing the process of collecting probe requests using the sensor} \label{datacollection_schematic}
\end{figure}

The next step after collecting data was to estimate the footfall or pedestrian activity from them. We identified the following major challanges which arise from our collection methodology.
\begin{enumerate}
	\item \textbf{Background noise} - since the extent to which Wi-Fi signals travel differs subject to various factors such as interference and humidity, it is close to impossible to restrict our data collection to a finite area of interest. This can lead to a signicant background noise at certain locations. E.g. a phone shop or a bus stop located next to the study area can increase the number of probe requests received by the sensor.
	\item \textbf{MAC randomisation} - The mobile devices in recent years have been using randomised 'local' MAC addresses for probe requests to protect the users from being tracked. This makes it impossible to tell if the probe requests are being sent by the same mobile device which is being stationed next to the sensor. This along with the previous problem can further increase the magnitude of error by several fold.
	\item \textbf{Mobile ownership} - Since the rate of mobile ownership can vary widely across geography and demography, we cannot assume that every mobile device translates to one pedestrian footfall. In addition to this, there is a long term overall increase in mobile ownership which may lead to the number of probe requests collected overtime. 
\end{enumerate}
We propose the following methods to tackle each of these challages.

\begin{figure}
	\centering
	\subfigure[Distribution of signal strengths showing the filtering of background noise]{
		\resizebox*{0.45\linewidth}{!}{\includegraphics{images/methodology_signalstrength.jpeg}}}\hspace{20pt}
	\subfigure[Clustering probe requests as nodes in a graph using increasing sequence numbers]{
	\resizebox*{0.45\linewidth}{!}{\includegraphics{images/methodology_clustering.jpeg}}}
	\caption{Schematic diagrams explaining the methods for filtering by signal strength and clustering using sequence numbers} \label{methodology_schematic}
\end{figure}

\subsection{Filtering with Signal Strength}
One of the clues that we can use to estimate the distance between the mobie device and the sensor is the strength of the signal received by the sensor. 
The obvious approach here is to try and establish a relationship between the signal strength and distance first and use this to filter out the unwanted probe requests.
This approach has numerous pitfalls and uncertainities since the decay of signal strength with distance is not always constant.
It varies with atmospheric conditions, presence of obstructions between the source and target, the nature of these obstructions and the strength (power level) of the source transponder.
This severely limits our ability to establishing a simple conversion between reported signal strength and distance.
There is a need for a method which takes in to account these variables across varous locations.

We hypothesise that in configurations where a specific source of background noise is at a constant distance, there must be a distinct break in the number of probe requests reporting signal strength corresponding to that distance.
For example, if there is a phone shop next to our sensor where hundreds of phones regularly send probe requests there should be a sharp rise of number of probe requests with reported singal strength corresponding to the distance between the sensor and the phone shop at any given set of conditions as shown in Figure \ref{methodology_schematic}.
We could identify these breaks in the data using traditional one diamensional classification algorithms such as Jenks natural breaks, k-means, quantile and hierarchical clustering, etc.
Since we are only looking for the break in the data and not for absolute values, the methodology should apply for all the variations due micro site conditions thus reducing the overall noise in the collected data.

@@

Since our primary unique identifier - MAC adddress, is being anonymised by new devces, we need to find other information present in the probe request for a unique identifier.
Obvious approach here is to establish a factor of randomisation and adjust the counts for these probe requests based on this factor.
We found this aproach not feasible, since the proportion of devices which randomise the MAC addresses increases over time.
There is also a wide variation in the frequency at which the devices randomise the MAC addresses and the method used for the process.
This leads us to look for a more generalisable approach which is independent of the device model.

From our initial analysis we found that OUI, length of the packet and sequence number of the packet being the most promising information to acheive this.
First we divide our dataset into sets of probe requests with randomised and non-randomised MAC addresses and keep the MAC address as the unique identifier for the latter set. 
For randomised ones we futher divide them in to sub categories based on their OUIand length of the packet.
Since the length tends to stay unique to specific models of devices we are left with the task of identifying the unique mobile devices from within these distinct models.

The proposed algorithm creates a graph where the probe requests represented the nodes, and links are created between them based on the following rules: 
\begin{itemize}
	\item A link could go only forward in time. 
	\item A link could go from low to high sequence numbers.
	\item A link could exist between nodes with a maximum time difference of $\alpha$ - time threshold.
	\item A link could exist between nodes with a maximum sequence number difference of $\beta$ - sequence threshold.
	\item A node could have only one incoming link and one outgoing link, which is the shortest of all such possible links.
\end{itemize}
The nodes were then classified based on the unique connected component they belong to as shown in Figure \ref{methodology_schematic}.
This classification was assigned as the unique identifier for the anonymised probe requests in the place of MAC address.
Though the recycling of sequence number after 4000 leads to multiple classifications reported on single device, the maginitude of error is greatly reduced.

\subsection{Calibrating with ground truth}
Since mobile phone ownership is an external uncertainty to our study and could arise from variety of spatio, temporal and demographic factors, we propose to solve this using external source of information. We hypothesize that an adjustment factor could be arrived at for each location of data collection, comparing the sensor based counts and ground truth and it can be used to adjust the data reliably to reflect the ground truth in absolute numbers for the future. This calibration can be carried over periodically and the frequecy of which will improve the quality of the estimation.

>>>>

The primary aim of this research is to enable us to collect a series of probe requests and process them into a usable pedestrian footfall count. 
We do this by using a Wi-Fi receiver to collect probe requests broadcast by mobile devices, filtering out the background noise and aggregating them based on the device that generated them.
In this section, we begin by looking at the characteristics of probe requests in detail, device a methodology to collect these probe requests in public areas, examine the systemic biases and uncertainties in the data collection method and device data processing methods to overcome these challenges.
Finally we compare the processed footfall counts to the ground truth recorded by primary surveys.

Probe requests are a special type of management packets broadcast by Wi-Fi enabled devices as part of the various functions such as scanning for available APs, quick geolocation by triangulation based known APs, etc.
These are broadcast by all Wi-Fi enabled devices regardless of the manufacturer, type or model of the devices though there is some variation on the frequency and the information transmitted through them.
In some cases, such as Android devices, these are broadcast even when the Wi-Fi functionality has been turned off by the user.
Thus these signals can be used to reliably identify the presence of Wi-Fi enabled mobile devices.
Being a first step of connection initiated by the mobile device, these packets have information regarding the characteristics of the mobile device itself. Some of the key information we can infer from these requests are,
\begin{enumerate}
	\item \textbf{Media Access Control (MAC) address} which is an unique identifier for the wireless hardware of the mobile device,
	\item \textbf{Sequence number} of the request for the mobile device to keep track of the responses,
	\item \textbf{Time stamp} at which the request was received by the AP,
	\item Total \textbf{length} of the request in number of bits, and 
	\item The \textbf{strength of the signal} which transmitted the request.
\end{enumerate}
The MAC address is the primary unique identifier for the mobile device.
It has two parts, first part is the Organisation Unique Identifier (OUI) which gives information about the manufacturer of the device and the second is unique to the device. In modern devices, to protect the users privacy, the MAC address can also be randomised (hence non unique) and is marked as such. Though sequence number and length of the packet are not strictly unique, we hypothesize that we can use them to estimate the number unique devices.

Data collection was done with the help of custom sensors built from modifying the hardware used in Smart street sensors \citep{sss2016} and updating them with custom software.
The sensor is essentially a Raspberry-Pi device with Wi-Fi and 3G modules.
It keeps the Wi-Fi module in `Monitor' mode and uses the open source software - Wireshark \citep{wireshark2} to passively collect all packets sent to `broadcast', marked with type as management' and subtype as `probe requests'.
The MAC address in these probe requests is anonymised at the device level using a cryptographic hashing algorithm and transmitted through 3G connection to a central database via web-sockets protocol, where it is stored in a PostgreSQL database for further analysis.
A overall schematic of the data collection and storage process is shown in Figure \ref{datacollection_schematic}.
The ground truth on number of pedestrian footfall was recorded using a custom mobile application - Clicker \citep{bala2018clicker}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{images/datacollection_schematic.jpeg}
	\caption{Schematic diagram showing the process of collecting and storing probe requests using the sensor} \label{datacollection_schematic}
\end{figure}

The next step after collecting data was to estimate the footfall or pedestrian activity from them. We identified the following potential uncertainties which arise from our collection methodology.
\begin{enumerate}
	\item \textbf{Background noise} - since the extent to which Wi-Fi signals travel differs subject to various factors such as interference and humidity, it is close to impossible to restrict our data collection to a finite area of interest. This can lead to a significant background noise at certain locations. E.g. a phone shop or a bus stop located next to the study area can increase the number of probe requests received by the sensor.
	\item \textbf{MAC randomisation} - The mobile devices in recent years have been using randomised 'local' MAC addresses for probe requests to protect the users from being tracked. This makes it impossible to tell if the probe requests are being sent by the same mobile device which is being stationed next to the sensor. This along with the previous problem can further increase the magnitude of error by several fold.
	\item \textbf{Mobile ownership} - Since the rate of mobile ownership can vary widely across geography and demography, we cannot assume that every mobile device translates to one pedestrian footfall. In addition to this, there is a long term overall increase in mobile ownership which may lead to the number of probe requests collected overtime. 
\end{enumerate}
We propose the following internal and external validation methods to tackle each of these uncertainties.

\begin{figure}
	\centering
	\subfigure[Distribution of signal strengths (dBm) showing the filtering of background noise]{
		\resizebox*{0.45\linewidth}{!}{\includegraphics{images/methodology_signalstrength.jpeg}}}\hspace{20pt}
	\subfigure[Clustering probe requests as nodes in a graph using increasing sequence numbers]{
	\resizebox*{0.45\linewidth}{!}{\includegraphics{images/methodology_clustering.jpeg}}}
	\caption{Schematic diagrams explaining the methods for filtering by signal strength and clustering using sequence numbers} \label{methodology_schematic}
\end{figure}

\subsection{Filtering with Signal Strength}
One of the clues that we can use to estimate the distance between the mobile device and the sensor is the strength of the signal received by the sensor. 
The obvious approach here is to try and establish a relationship between the signal strength and distance first and use this to filter out the unwanted probe requests.
This approach was found not to be feasible since the decay of signal strength with distance is not always constant.
It varies with atmospheric conditions, presence of obstructions between the source and target, the nature of these obstructions and the strength (power level) of the source.
This severely limits our ability in establishing a simple conversion between reported signal strength and distance.
There is a need for a method which takes in to account these variables across various locations.

We hypothesise that in configurations where a specific source of background noise is at a constant distance, there must be a distinct pattern in the number of probe requests reporting signal strength corresponding to that distance.
For example, if there is a phone shop next to our sensor where hundreds of phones regularly send probe requests there should be a sharp rise of number of probe requests with reported signal strength corresponding to the distance between the sensor and the phone shop irrespective of the local conditions as shown in Figure \ref{methodology_schematic}.
We could identify these breaks in the data using traditional one dimensional classification algorithms such as `jenks natural breaks', `k-means', `quantile' and `hierarchical clustering', etc.
Since we are only looking for the break in the data and not for absolute values, the methodology should apply for all the variations due micro site conditions thus reducing the overall noise in the collected data.

@@

Since our primary unique identifier - MAC address, is being anonymised by new devices, we need to find other information present in the probe request for use as a unique identifier.
Obvious approach here is to establish a factor of randomisation and adjust the counts for these probe requests based on this factor.
We found this approach to be not feasible since the proportion of devices which randomise the MAC addresses increases over time.
There is also a wide variation in the frequency at which the devices randomise the MAC addresses and the method used for the process.
This lead us to look for a more generalisable approach which is independent of the device model.

From our initial analysis we found that OUI, length of the packet and sequence number of the packet being the most promising information to achieve this.
First we divide our dataset into sets of probe requests with randomised and non-randomised MAC addresses and keep the MAC address as the unique identifier for the latter set. 
For randomised ones we further divide them in to sub categories based on their OUI and length of the packet.
Since the length tends to stay unique to specific models of devices we are left with the task of identifying the unique mobile devices from within these sets.

The proposed algorithm creates a graph where the probe requests represented the nodes, and links are created between them based on the following rules: 
\begin{itemize}
	\item A link could go only forward in time. 
	\item A link could go from low to high sequence numbers.
	\item A link could exist between nodes with a maximum time difference of $\alpha$ - time threshold.
	\item A link could exist between nodes with a maximum sequence number difference of $\beta$ - sequence threshold.
	\item A node could have only one incoming link and one outgoing link, which is the shortest of all such possible links in terms of both time and sequence number.
\end{itemize}
The nodes were then classified based on the unique connected component they belong to as shown in Figure \ref{methodology_schematic}.
This classification was assigned as the unique identifier for the anonymised probe requests in the place of MAC address.
Though the recycling of sequence number after 4000 leads to multiple classifications reported on single device, the magnitude of error is greatly reduced.

\subsection{Calibrating with Ground Truth}
Since mobile device ownership is an external uncertainty to our study and could arise from variety of spatio - temporal and demographic factors, we propose to solve this using external source of information.
We hypothesize that an adjustment factor could be arrived at for each location of data collection, comparing the sensor based counts and ground truth and it can be used to adjust the data reliably to reflect the ground truth in absolute numbers for the future.
This calibration can be carried over periodically and the frequency to improve the quality of the estimation.

### 2018-11-14 paper/sections/pilot_study.tex
wednesday modifications

To start we designed a small pilot study to validate the classification and clustering methodology against the scale and complexity of data collected on a open public area such as a retail highstreet.
We also aim to find the algorithm which is best suited for classification of signal strengths to filter out the background noise.
The data was collected at Oxford Street, London on 20 December 2017 from 12:30 to 13:00 hrs, where Wi-Fi probe requests were collected using the sensor described in Chapter \ref{methodology} and pedestrian footfall was manually recorded using the Android app \citep{bala2018clicker}.
Being located at one of the busiest retail locations in the United Kingdom, the Wi-Fi sensor captured approximately 60,000 probe requests during the half hour period, and 3,722 people were recorded manually walking on the sidewalk during that time.

\begin{figure}
	\centering
	\subfigure[Clustering probe requests based on increasing sequence numbers present in them.]{
		\resizebox*{0.45\linewidth}{!}{\includegraphics{images/pilot_clustering.jpeg}}}\hspace{20pt}
	\subfigure[Comparision of sensor based counts with the groud truth.]{
	\resizebox*{0.45\linewidth}{!}{\includegraphics{images/pilot_comparison.jpeg}}}
	\caption{The results of the pilot study demonstrating the validity of the deviced methodology.} \label{pilot_results}
\end{figure}

As a first step we just aggregated the probe requests by their MAC address for every minute to generate a minute by minute count of the number of people around the sensor assuming each MAC address corresponds to a mobile device and hence a pedestrian.
We then compare this preliminary `footfall' count to the actual number of pedestrian recorded manually to check for the robustness. We use Mean Absolute Percentage Error (MAPE) measure of robustness of the count since it provides a simple and quick measurement while the street conditions ensure that there is no `zero records' issue.
We find that the MAPE in the raw counts compared to the ground truth is around 425\%.
This suggests the presence of large amount of noise in the data which might be generated due to sources of uncertainties discussed in Chapter \ref{methodology} thus demonstrating the need for filtering the data.

We then classified the probe requests as ones with "high signal strength" and "low signal strength" using various classification algorithm such as k-means, quantile, heirarchical clusters, fisher and jeanks.
We found that while heirarchical clustering and jenks were resource intensive, kmeans gave quickest results with the lowest MAPE closely followed by quantile.
The cut-off point or threshold for the collected data with which we could classify them as high and low was -71 dBm.
We then removed all the probe requests which reported `low signal strength' and repeated the same aggregation process as before to produce footfall count. 
This process resulted in a footfall count with a net MAPE of 30\%.
Though the results are encouraging we are still not completely confident that our filtering process is indeed removing noise or has any correlation the configuration of sensor or position of the mobile devices.
These need to be addressed with a larger survey with multiple location of varying orientations.

The next challenge was to identify probe requests which are generated by the same device irrespective of MAC randomisation process.
We use the algorithm defined in Chapter \ref{methodology} and assign a unique identifier or signature to each probe request independent of the MAC address. 
Since there is neither prior research nor documentation on a universal behaviour of phones in randomising their MAC addresses, we use the stationary device - the one used for manual counting as a reference and find out the suitable time threshold, $\alpha$ and threshold for sequence numbers, $\beta$ to be 16s and 60 respectively via trial and error.
This process is done on top the filtering done based on signal strength and only for the probe requests with randomised MAC addresses.
Figure \ref{pilot_results} shows the results of this clustering process on a small set of randomised probe requests.
The probe request with different randomised MAC address is shown by the colored points and the line joining them shows the ones belonging to the same cluster hence expected to be generated by the same device.
We finally aggregate the probe requests as before but with the device signature rather than just MAC addresses this results in a footfall count with a MAPE of -18\%. 
A comparision of minute by minute counts resulting from different filtering processes along with the ground truth is shown in Figure \ref{pilot_results} showing the promising effectiveness of the methods.

>>>>

To start we designed a small pilot study to validate the classification and clustering methodology against the scale and complexity of data collected on a open public area such as a retail high street.
We also aim to find the algorithm which is best suited for classification of signal strengths to filter out the background noise.
The data was collected at Oxford Street, London on 20 December 2017 from 12:30 to 13:00 hrs, where Wi-Fi probe requests were collected using the sensor described in Chapter \ref{methodology} and pedestrian footfall was manually recorded using the Android app \citep{bala2018clicker}.
Being located at one of the busiest retail locations in the United Kingdom, the Wi-Fi sensor captured approximately 60,000 probe requests during the half hour period, and 3,722 people were recorded manually walking on the sidewalk during that time.

As a first step we just aggregated the probe requests by their MAC address for every minute to generate a minute by minute count of the number of people around the sensor assuming each MAC address corresponds to a mobile device and hence a pedestrian.
We then compare this preliminary `footfall' count to the actual number of pedestrian recorded manually to check for the robustness. We use Mean Absolute Percentage Error (MAPE) measure of robustness of the count since it provides a simple and quick measurement while the street conditions ensure that there are no intervals without any footfall.
We find that the MAPE in the raw counts compared to the ground truth is around 425\%.
This suggests the presence of large amount of noise in the data which might be generated due to sources of uncertainties discussed in Chapter \ref{methodology} thus demonstrating the need for filtering the data.

We then classified the probe requests as ones with "high signal strength" and "low signal strength" using various classification algorithm such as k-means, quantile, hierarchical clusters, fisher and jenks.
We found that while hierarchical clustering and jenks were resource intensive, k-means gave quickest results with the lowest MAPE closely followed by quantile.
The cut-off point or threshold for the collected data with which we could classify them as high and low was -71 dBm.
We then removed all the probe requests which reported `low signal strength' and repeated the same aggregation process as before to produce footfall count. 
This process resulted in a footfall count with a net MAPE of 30\%.
Though the results are encouraging we are still not completely confident that our filtering process is indeed removing noise or has any correlation the configuration of sensor or position of the mobile devices.
These need to be addressed with a larger survey with multiple location of varying orientations.

\begin{figure}
	\centering
	\subfigure[Clustering probe requests based on increasing sequence numbers present in them.]{
		\resizebox*{0.45\linewidth}{!}{\includegraphics{images/pilot_clustering.jpeg}}}\hspace{20pt}
	\subfigure[Comparison of sensor based counts with the ground truth.]{
	\resizebox*{0.45\linewidth}{!}{\includegraphics{images/pilot_comparison.jpeg}}}
	\caption{The results of the pilot study demonstrating the validity of the methodology.} \label{pilot_results}
\end{figure}

The next challenge was to identify probe requests which are generated by the same device irrespective of MAC randomisation process.
We use the algorithm defined in Chapter \ref{methodology} and assign a unique identifier or signature to each probe request independent of the MAC address. 
Since there is neither prior research nor documentation on a universal behaviour of phones in randomising their MAC addresses, we use the stationary device - the one used for manual counting as a reference and find out the suitable time threshold, $\alpha$ and threshold for sequence numbers, $\beta$ to be 16s and 60 respectively via trial and error.
This process is done on top the filtering done based on signal strength and only for the probe requests with randomised MAC addresses.
Figure \ref{pilot_results} shows the results of this clustering process on a small set of randomised probe requests.
The probe request with different randomised MAC address is shown by the colored points and the line joining them shows the ones belonging to the same cluster hence expected to be generated by the same device.
We finally aggregate the probe requests as before but with the device signature rather than just MAC addresses this results in a footfall count with a MAPE of -18\%. 
A comparison of minute by minute counts resulting from different filtering processes along with the ground truth is shown in Figure \ref{pilot_results} showing the promising effectiveness of the methods.

### 2018-11-14 paper/sections/previous_work.tex
wednesday modifications

Though Wi-Fi is a `location-less' technology, there are reliable methods to triangulate the location of Wi-Fi enabled mobile devices from the locations of known access points(AP) and the signal strength reported by them \citep{he2003range, moore2004robust, lamarca2005place}.
This can overcome the usual shortcoming of GPS, which struggles for precision and accuracy in indoor and densely built environments \citep{zarim2006,kawaguchi2009wifi, xi2010locating}. 
Utilising this, we can easily and quickly estimate trajectories of the mobile devices \citep{Sorensenlocation,musa2012tracking} which can be used similar to the GPS trajectories to understand individual travel patterns \citep{reki2007,Sap2015}, crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular \citep{lu2010vehicle} and pedestrian movement \citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}.
Such data can also be used in transportation planning and management to estimate travel time \citep{musa2011wiflow} and real time traffic monitoring \citep{abbott2013empirical}.
Using techniques demonstrated by \cite{franklin2006passive} and \cite{pang2007802} along with information present in the probe requests one even model interactions between the users \citep{cheng2012inferring,barbera2013signals,cunche2014know,cunche2014linking} such as predicting which of them are most likely to meet again \citep{cunche2012know}.
Using the semantic information present in these probe requests it even is possible to understand the nature of population at a large scale \citep{di2016mind}.

Though extensive research has been carried out on this subject with feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the increasing attempt by mobile phone manufacturers to protect their users’ privacy by anonymising the globally identifiable portion of the probe requests, \citep{green2008}.
Various methods have been devised to overcome this anonymisation process such as estimating the device model information from a known dataset of manufacturers and device behaviours \citep{martin2016decomposition}; Scrambler attack using a small part of the physical layer specification for Wi-Fi \citep{vo2016,bloessl2015scrambler}; and timing attack where the packet sequence information along with information elements present in the probe request frame is used \citep{matte2016,cheng2016can}.
A combination of these methodologies has been proven to produce de-anonymised globally unique device information \citep{vanhoef2016, martin2017}.
These approaches usually result in serious risk of infringement of the privacy of the users of the mobile devices by revealing their identifiable personal information.

There is a clear gap in research for exploring methodologies for estimating the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original device information.
Such technique has various applications such as uncovering the urban wireless landscape \citep{rose2010mapping}, revealing human activity at large scales \citep{qin2013discovering}, estimating pedestrian numbers in crowds \citep{schauer2014estimating,fukuzaki2015statistical} and even counting people in hyper local scales such as queues \citep{wang2013measuring}
With enough infrastructure to collect such information we can even aim to generate a real-time census of the city \citep{konto2017}.
With this background we set out to device and implement a methodology to reliably estimate human activity such as pedestrian footfall from Wi-Fi probe requests without infringing the privacy of the users involved.

>>>>

Though, unlike GPS, the location of the Wi-Fi enabled mobile device cannot be directly inferred from Wi-Fi, there are reliable methods to triangulate the location of mobile devices from the locations of known access points(AP) and the signal strength reported by them \citep{he2003range, moore2004robust, lamarca2005place}.
This can overcome the usual shortcoming of GPS, which struggles for precision and accuracy in indoor and densely built environments \citep{zarim2006,kawaguchi2009wifi, xi2010locating}. 
Utilising this, we can easily and quickly estimate trajectories of the mobile devices \citep{Sorensenlocation,musa2012tracking} which can be used similar to the GPS trajectories to understand individual travel patterns \citep{reki2007,Sap2015}, crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular \citep{lu2010vehicle} and pedestrian movement \citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}.
Such data can also be used in transportation planning and management to estimate travel time \citep{musa2011wiflow} and real time traffic monitoring \citep{abbott2013empirical}.
Using techniques demonstrated by \cite{franklin2006passive} and \cite{pang2007802} along with information present in the probe requests one even model interactions between the users \citep{cheng2012inferring,barbera2013signals,cunche2014know,cunche2014linking} such as predicting which of them are most likely to meet again \citep{cunche2012know}.
Using the semantic information present in these probe requests it even is possible to understand the nature of population at a large scale \citep{di2016mind}.

Though extensive research has been carried out on this subject with feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the increasing attempt by mobile phone manufacturers to protect their users’ privacy by anonymising the globally identifiable portion of the probe requests, \citep{green2008}.
Various methods have been devised to overcome this anonymisation process such as estimating the device model information from a known dataset of manufacturers and device behaviours \citep{martin2016decomposition}; Scrambler attack using a small part of the physical layer specification for Wi-Fi \citep{vo2016,bloessl2015scrambler}; and timing attack where the packet sequence information along with information elements present in the probe request frame is used \citep{matte2016,cheng2016can}.
A combination of these methodologies has been proven to produce de-anonymised globally unique device information \citep{vanhoef2016, martin2017}.
These approaches usually result in serious risk of breach of privacy of the users of the mobile devices by revealing their identifiable personal information.

There is a clear gap in research for exploring methodologies for estimating the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original device information.
Such technique has various applications such as uncovering the urban wireless landscape \citep{rose2010mapping}, revealing human activity at large scales \citep{qin2013discovering}, estimating pedestrian numbers in crowds \citep{schauer2014estimating,fukuzaki2015statistical} and even counting people in hyper local scales such as queues \citep{wang2013measuring}
With enough infrastructure to collect such information we can even aim to generate a real-time census of the city \citep{konto2017}.
With this background we set out to device and implement a methodology to reliably estimate human activity such as pedestrian footfall from Wi-Fi probe requests without risking the breach of privacy of the users involved.
